### 2014-11-12 contract-language.tex
Incorporated new edits
Back to 11 pages

\begin{figure}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

\begin{itemize}
\setlength{\itemsep}{2pt}

\item \textbf{Eventual consistency}: Eventually consistent operations can
	be satisfied as long as the client can reach at least one replica. In the
	bank account example, \cf{deposit} is an eventually consistent operation.
	While eventually consistent data stores typically offer \emph{basic} eventual
	consistency with all possible anomalies, we assume that our store provides
	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
	store always exposes a \emph{causal cut} of the updates. This semantics can
	be formally captured in terms of the following contract definition:

@@

\newcommand{\DDe}[1]{#1}
\begin{figure}
\begin{smathpar}
\begin{array}{c}
\vspace{3mm}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  \qquad

\RuleTwo
{\DDe{\cv} \le \DDe{\ecc}}
{{\sf EventuallyConsistent}(\cv)} \\

\vspace{3mm}
\RuleTwo
{\DDe{\cv} \not\le \DDe{\ecc}
\quad \DDe{\cv} \le \DDe{\ccc}}
{{\sf CausallyConsistent}(\cv)} \qquad

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\vspace{-5mm}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{itemize}
\setlength{\itemsep}{2pt}
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

>>>>

\begin{figure}
\vspace{-7pt}
\begin{smathpar}
\begin{array}{lclcl}
\multicolumn{5}{l}{
  {\eff} \in \mathtt{Effect} \qquad
  {\cv} \in \mathtt{Contract} \qquad
  \set{\eff} \in \mathtt{Effect\; Set}
}\\
\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
%\sameobjZ		&     &  & \\
{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
\end{array}
\end{smathpar}

@@

\begin{itemize}[noitemsep]

\item \textbf{Eventual consistency}: Eventually consistent operations can
	be satisfied as long as the client can reach at least one replica. In the
	bank account example, \cf{deposit} is an eventually consistent operation.
	While an \ecds typically offer \emph{basic} eventual consistency with all
	possible anomalies, we assume that our store provides stronger semantics that
	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
	\emph{causal cut} of the updates. This semantics can be formally captured in
	terms of the following contract definition:

@@

\newcommand{\DDe}[1]{#1}
\begin{figure}
\vspace{-10pt}
\begin{smathpar}
\begin{array}{cc}
\RuleTwo
{\DDe{\cv} \le \DDe{\scc}}
{{\sf WellFormed}(\cv)}  &

\RuleTwo
{\DDe{\cv} \le \DDe{\ecc}}
{{\sf EventuallyConsistent}(\cv)} \\ \\

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ecc}
\quad \DDe{\cv} \le \DDe{\ccc}}
{{\sf CausallyConsistent}(\cv)} &

\RuleTwo
{\DDe{\cv} \not\le \DDe{\ccc}
\quad \DDe{\cv} \le \DDe{\scc}}
{{\sf StronglyConsistent}(\cv)}
\end{array}
\end{smathpar}
\vspace{-3mm}
\caption{Contract classification.}
\label{sem:classify}
\end{figure}

@@

\begin{itemize}[noitemsep]
\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
\item Visibility only relates actions on the same object:
	\begin{itemize}
	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
	\end{itemize}
\item Session order is a transitive relation:
	\begin{itemize}
	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
	\end{itemize}
\item Same object is an equivalence relation:
	\begin{itemize}
	\item $\small \forall a.~\sameobj{a}{a}$.
	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
	\end{itemize}
\end{itemize}
\end{definition}

### 2014-11-12 evaluation.tex
Back to 11 pages

\begin{itemize}
\setlength{\itemsep}{2pt}
\item \textbf{LWW register}: A last-write-wins register that provides read
	and write operations, where the read returns the value of the latest write.

@@

\item \textbf{Bank account}: Our running example, with savings and checking
  accounts.

@@

For our performance evaluation, we deploy \name applications in
\emph{clusters}, where each cluster is composed of five fully replicated
Cassandra replicas within the same datacenter. We instantiate one shim layer
node co-located on the same VM as a Cassandra replica. Clients are instantiated
within the same data center as the store, and run transactions. We deploy each
cluster node on an \cf{c3.4xlarge} Amazon EC2 instance.  Our shim layer nodes
are multi-threaded, and we allocate 8 CPUs (out of 16 available) for each shim
layer node. The clients also run on \cf{c2.4xlarge} instances. We call this a
\cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
instantiate 2 clusters, each with five nodes, and place the clusters on US-east
(Virginia) and US-west (Oregon) locations. The average inter-region latency was
85ms.

@@

The lines marked EC and CC correspond to all operations (including
\cf{withdraw}) being assigned EC and CC consistency levels. These levels
compromise correctness as \cf{withdraw} has to be an SC operation. The SC line
corresponds to a configuration where all operations are strongly consistent;
this ensures application correctness, at the cost of performance. \name
corresponds to our implementation, which classifies operations based on their
contract specifications.  Both \name and SC ensure correctness. With 512
clients, the \name implementation was within 41\% of the latency and 18\% of
the throughput of EC, whereas SC operations had 162\% higher latency and 52\%
lower throughput than EC operations. Observe that there is a point in each case
after which the latency increases while the throughput decreases; these
correspond to points where the store becomes saturated with client requests.

In a \cf{2DC} configuration (not shown here), the average latency of SC
operations with 512 clients increased by 9.4$\times$ due to the cost of
geo-distributed coordination, whereas \name operations were only 2.2$\times$
slower, mainly due to the increased cost of \cf{withdraw} operations.
Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
the same, illustrating the benefit of fine-grained contract classification.

>>>>

\begin{itemize}[noitemsep]
\item \textbf{LWW register}: A last-write-wins register that provides read
	and write operations, where the read returns the value of the latest write.

@@

\item \textbf{Bank account}: Our running example.

@@

For our performance evaluation, we deploy \name applications in
\emph{clusters}, where each cluster is composed of five fully replicated
Cassandra replicas within the same datacenter. We instantiate one shim layer
node co-located on the same VM as a Cassandra replica. Clients are instantiated
within the same data center as the store, and run transactions. We deploy each
cluster and client node on an \cf{c3.4xlarge} Amazon EC2 instance. We call this
a \cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
instantiate 2 clusters, each with five nodes, and place the clusters on US-east
(Virginia) and US-west (Oregon) locations. The average inter-region latency was
85ms.

@@

The lines marked EC and CC correspond to all operations (including
\cf{withdraw}) being assigned EC and CC consistency levels. These levels
compromise correctness as \cf{withdraw} has to be an SC operation. The SC line
corresponds to a configuration where all operations are strongly consistent;
this ensures application correctness, at the cost of performance. \name
corresponds to our implementation, which classifies operations based on their
contract specifications. With 512 clients, the \name implementation was within
41\% of the latency and 18\% of the throughput of EC, whereas SC operations had
162\% higher latency and 52\% lower throughput than EC operations. Observe that
there is a point in each case after which the latency increases while the
throughput decreases; these correspond to points where the store becomes
saturated with client requests. In a \cf{2DC} configuration (not shown here),
the average latency of SC operations with 512 clients increased by 9.4$\times$
due to the cost of geo-distributed coordination, whereas \name operations were
only 2.2$\times$ slower, mainly due to the increased cost of \cf{withdraw}
operations. Importantly, the latency of \cf{getBalance} and \cf{deposit}
remained almost the same, illustrating the benefit of fine-grained contract
classification.

### 2014-11-12 model.tex
Incorporated new edits
Back to 11 pages

\begin{figure}
\centering
\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
\caption{\name system model.}
\label{fig:sysmod}
\end{figure}

@@

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\small \vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$,
where $\small \visZ$ is the visibility relation between effects. Visibility is
an irreflexive and asymmetric relation, and only relates effects produced by
operations on the same object. Executing a read-only operation is similar
except that no new effects are produced. The effect added to a particular
replica is asynchronously sent to other replicas, and eventually merged into
all other replicas. Observe that this model does not assume a particular
resolution strategy for concurrent conflicting updates, and instead preserves
\emph{every} update. Conflicts are resolved when an operation reduces over the
set of effects on an object at a particular replica.

@@

This model admits all the inconsistencies associated with eventual consistency.
The goal of this work is to identify the precise consistency level for each
operation such that application-level constraints are not violated. In the next
section, we will concretely describe the challenges associated with
constructing a consistent bank account on top of an eventually consistent data
store. Subsequently, we will illustrate how our contract and specification
language, armed with the primitive relations $\small \visZ$, $\small \soZ$,
$\small \sameobjZ$ and $\small \operZ$, mitigates these challenges.

>>>>

\begin{figure}
\centering
\includegraphics[width=0.7\columnwidth]{Figures/SystemModel}
\caption{\name system model.}
\label{fig:sysmod}
\end{figure}

@@

When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
replica 1, it simply \emph{reduces} over the current set of effects at that
replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
appended to the state of $x$ at replica 1. Thus, every operation is evaluated
over a \emph{snapshot} of the state of the object on which it is invoked. In
this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
written logically as $\small \vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$,
where $\small \visZ$ is the visibility relation between effects. Visibility is
an irreflexive and asymmetric relation, and only relates effects produced by
operations on the same object. Executing a read-only operation is similar
except that no new effects are produced. The effect added to a particular
replica is asynchronously sent to other replicas, and eventually merged into
all other replicas. Observe that this model does not assume a particular
resolution strategy for concurrent conflicting updates, and instead preserves
\emph{every} update. Update conflicts are resolved when an operation reduces
over the set of effects on an object at a particular replica.

@@

This model admits all the inconsistencies associated with eventual consistency.
The goal of this work is to identify the precise consistency level for each
operation such that application-level constraints are not violated. In the next
section, we will concretely describe the challenges associated with
constructing a consistent bank account on top of an \ecds. Subsequently, we
will illustrate how our contract and specification language, armed with the
primitive relations $\small \visZ$, $\small \soZ$, $\small \sameobjZ$ and
$\small \operZ$, mitigates these challenges.

### 2014-11-12 motivation.tex
Back to 11 pages

Consider how we might implement a highly available bank account on top of an
eventually consistent data store, with the \emph{integrity} constraint that the
balance must be non-negative. We begin by implementing a bank account
replicated data type (RDT) in \name, and then describe the mechanisms to obtain
the desired correctness guarantees.

@@

The datatype \cf{Acc} represents the effect type for the bank account. The
function \cf{sum} returns the sum of elements in the list, and \cf{sel1}
returns the first element of a tuple. For each operation, \cf{hist} is a
\emph{snapshot} of the state of the object at some replica. In this sense,
every operation on the RDT is atomic, and thus permitting sequential reasoning
for implementing eventually consistent data types. Here, \cf{getBalance} is a
read-only operation, \cf{deposit} always emits an effect, and \cf{withdraw}
only emits an effect if there is sufficient balance in the account. We have
implemented a large corpus of RDTs for realistic benchmarks including shopping
carts, auction and micro-blogging sites in few tens of lines of code.

>>>>

Consider how we might implement a highly available bank account on top of an
\ecds, with the \emph{integrity} constraint that the balance must be
non-negative. We begin by implementing a bank account replicated data type
(RDT) in \name, and then describe the mechanisms to obtain the desired
correctness guarantees.

@@

The datatype \cf{Acc} represents the effect type for the bank account. The
function \cf{sum} returns the sum of elements in the list, and \cf{sel1}
returns the first element of a tuple. For each operation, \cf{hist} is a
\emph{snapshot} of the state of the object at some replica. In this sense,
every operation on the RDT is atomic, and thus permitting sequential reasoning
for implementing the RDTs. Here, \cf{getBalance} is a read-only operation,
\cf{deposit} always emits an effect, and \cf{withdraw} only emits an effect if
there is sufficient balance in the account. We have implemented a large corpus
of RDTs for realistic benchmarks including shopping carts, auction and
micro-blogging sites in few tens of lines of code.

### 2014-11-12 transactions.tex
Applied changes from paper to techrep
Back to 11 pages

\begin{figure}
\centering
\subfigure[Atomicity]{\label{fig:txn_atomicity}\includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}}
\hfill
\subfigure[Monotonic Atomic View]{\label{fig:txn_mav}\includegraphics[width=0.25\columnwidth]{Figures/TxnMAV}}
\hfill
\subfigure[Repeatable Read]{\label{fig:txn_rr}\includegraphics[width=0.25\columnwidth]{Figures/TxnRR}}
\caption{Semantics of transaction contracts. $x$ and $y$ are distinct objects.
The dotted line represents the visibility requested by the contracts.}
\label{fig:transaction}
\end{figure}

@@

A transaction with ANSI RC semantics only witnesses committed operations. Let
us assume that the store buffers updates until all the updates from the
transaction are available at a replica. If the transaction commits, the
buffered updates are made visible. Otherwise, the buffered updates are
discarded. RC does not entail any further isolation guarantees. Hence, a store
implementing RC does not require inter-replica coordination. We can express RC
as follows:

>>>>

\begin{figure}
\vspace{-2.3mm}
\centering
\subfigure[Atomicity]{\label{fig:txn_atomicity}\includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}}
\hfill
\subfigure[Monotonic Atomic View]{\label{fig:txn_mav}\includegraphics[width=0.25\columnwidth]{Figures/TxnMAV}}
\hfill
\subfigure[Repeatable Read]{\label{fig:txn_rr}\includegraphics[width=0.25\columnwidth]{Figures/TxnRR}}
\caption{Semantics of transaction contracts. $x$ and $y$ are distinct objects.
The dotted line represents the visibility requested by the contracts.}
\label{fig:transaction}
\end{figure}

@@

A transaction with ANSI RC semantics only witnesses committed operations. Let
us assume that a replica will buffer transactional updates until all the
updates from the same transaction are available at that replica. Once all the
updates from a transaction are available, the buffered updates are made visible
to subsequent client requests. This ensure atomicity of transactions.
Importantly, RC does not entail any other guarantees. As a result, a store
implementing RC does not require inter-replica coordination. We can express RC
as follows:
